---
title: "Penguin Flipper Length ANOVA"
author: "Terra Dressler"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")


```
Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA...
-independnet onservations
-equal variance (Levene's test)
-normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(common_name = case_when(species == "Chinstrap penguin (Pygoscelis antarctica)
" ~ "Chinstrap", species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie")) %>% 
  filter(sex == "FEMALE")




```
Exploratory graphs and Levene's test for equal variances
```{r}
hists <- ggplot(penguins2, aes(x= flipper_length))+ geom_histogram(aes(fill = "common_name")) + facet_wrap(~common_name, scale = "free")

hists

qqs<- ggplot(penguins2, aes(sample = flipper_length)) + geom_qq(aes(color = "common_name")) + facet_wrap(~common_name, scale = "free")

qqs

#Based on qq plots and histograms, each group appears normally distributed. Also, n>30 for all groups, so means will be normally distributed by CLT.
```

Evaluate variances and prform a Levene's test for equal variances
```{r}
#If the largest sample vriances is <4x greater than the smallest sample variance, then these are "close enough" to consider variances equal.

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(variance = var(flipper_length))
variances

#The largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)
#Levene's test
#leveneTest function in the "car" package

levene_penguin <-leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

#H0 no difference in variances
#HA variances are not equal

#There is no significant difference in variances across species.
```
One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}
#H0 means across all groups are equal
#HA at least 2 group means differ significantly 

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

#Post-hoc testing by Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#aov test funtion does allow for in-line  references
```
Mean flipper lengths of feale penguins differed significantly for all secies studied (one-way anovam F(df)= statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's $\alpha$ = 0.05)
